Improving the Image Acquisition Rate of an Atomic Force Microscope Through Compressive Sampling

### 2019-06-17 latex/afm-cs-main.tex
Tweak figures. Tighten up prose.

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim to reduce
  the time to acquire an image by reducing the number of measurements needed
  while still maintaining image quality. In this paper, we describe a hardware
  implementation and demonstration of one such approach based on the use of
  collections of short scans known as $\mu$--paths. Using a commercial AFM, we
  acquire data on a grating sample.
  Reconstructions are made from these data using
  %three approaches: (1) inpainting, an interpolation technique that
  %diffuses information from sampled locations into unsampled regions, (2)
  %basis-pursuit, a compressive sensing-based algorithm that seeks to minimize a
  %measure of sparsity in the underlying image, and (3)
  a new variant of basis
  pursuit, Basis Pursuit with Vertical Variation (BPVV), that is designed to reduce
  artifacts arising from the sampling pattern. The quality of the resulting
  images is compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio and the
  structured similarity index metric
  and a new metric we call the relative damage index.
  \todo{Re-write abstract, especially conclusions}
  These experiments demonstrate that at slow
  scan rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction) and that at
  high rates, the undersampling scheme produces higher quality images than
  raster scanning.
\end{abstract}

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of imaging
sample surface topography, material characteristics, and other surface
properties, with nanometer scale precision. AFMs acquire information about the sample by monitoring the deflection of the cantilever caused by the tip-sample interaction force. In general, 
%AFMs acquire information about the sample
%through a variety of imaging modes, all of which rely on the deflection of the
%cantilever caused by the tip-sample interaction force. While the details of each
%of the modes are quite different, in general, 
feedback control is used to hold
the deflection signal constant and information about the property of interest is
inferred from the applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial resolution, and
ability to image in vacuum, air, and liquid, AFM is widely used in a variety of
disciplines, including physics, biology and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}

@@

% % ------------------ TKTKTKT ------------------
% \todo{Strike this?}
% First, we demonstrate
% performance using both randomly-placed $\mu-$paths and designed $\mu-$paths that
% are optimized to minimize the expected reconstruction error given some initial
% estimate of the frequency structure of the sample. These schemes are described
% in Sec. \ref{sec:samplingScheme}.
% % --------------------------
We also introduce a new reconstruction
algorithm which is designed to minimize the artifacts arising from the structure
of the $\mu$-path scanning pattern. This algorithm is described and compared to
existing reconstruction methods through simulations in Sec.
\ref{sec:reconstructionMethods}.
Section \ref{sec:control} describes improvements to the control algorithms compared to our preliminary implementation in ~\cite{braker_hardware_2018}.
We then demonstrate the entire scheme from
sampling through reconstruction using experiments on a commercial AFM (Agilent
5500). The experimental setup and implementation details are described in Secs.
\ref{sec:experimentalSetup} and \ref{sec:implementation}\footnote{The software
  portion of the implementation and image reconstruction described in this paper
  can be found at \url{https://github.com/rabraker/AFM_CS}}. Experimental results are provided in Sec.
\ref{sec:results}. Finally, we provide a few concluding remarks in
Sec. \ref{sec:conclusions}.

@@

Although BPDN is effective in the general setting, using it for reconstruction
from horizontal $\mu$-path samples often yields artifacts in the vertical
direction (that is, the direction orthogonal to the $\mu$-path scans) leading to
strong discontinuities in the image. These artifacts grow more prominent as the
length of the $\mu$-paths is increased \cite{maxwell2014compressed}, because these
longer paths lead to increased mutual coherence between the measurement matrix
$R$ and the sparsity basis $U$. However, longer $\mu$-paths means fewer
$\mu$-paths and thus shorter scan times.
\noindent 
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------ Row  2 --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first column)
    Original raster-scanned image. (second column) BPDN reconstruction from
    random, 40 pixel long $\mu$-paths with 25\% sampling. (third column) BPVV
    reconstruction from the same sub-sampled data. (remaining columns)
    Corresponding details from the red boxes indicated in the raster image. The
    results show that BPVV reduces the artifacts arising from the horizontal
    scans of the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

To mitigate these affects, we propose a new variant of BPDN that we term basis pursuit with
vertical variation (BPVV). BPVV adds a vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to
\todo{It has been my observation that some benefit is gained by including $D_h$, but weighting it less than $D_v$. Not sure if we should leave it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                         % \quad W = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T \nonumber
\end{align}
where
\begin{equation*}
W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp., ${D_v x=\text{vec}(\nabla_v X)}$) representing a 
discrete gradient of a matrix in the vertical (resp., horizontal) direction (see, e.g., \cite[Section 6.1,]{becker_nesta_2011}). 
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
%   (\nabla_h X)_{i,j} = 
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp., $\left \| D_h x \right \|_1$) is the total variation (TV) of the signal in the vertical (resp., horizontal) direction.
The parameters $\alpha$ and $\beta$ are weighting parameters.
The description of BPVV in \eqref{op:bp_tvW} can be interpreted as changing our assumption of sparsity in the DCT basis to assuming sparsity in an overcomplete dictionary \cite{candes_redundant_2011}.
\todo{I believe this is correct: Sean/Yufan can you confirm or reject?}
When written as \eqref{op:bp_tvW}, the problem is in a form that may be solved with NESTA~\cite{becker_nesta_2011}.

% For completeness, we give an overview here.
% To that end, we re-write the cost $f(x)$ in \eqref{op:bp_tvW} as
% \begin{equation*}
%   f(x) = \max_{u\in Q_d} \left\langle u, W^Tx \right\rangle,\quad Q_d = \{u:~||u||_{\infty} \leq 1\}.
% \end{equation*}

% One of the core features of the NESTA algorithm is to replace the non-smooth functional $f(x)$ with a smoothed approximation
% \begin{equation}
%   f_{\mu} = \max_{u\in Q_d} \left\langle u,
%     W^T x \right\rangle - \frac{\mu}{2}||u||_2^2. \label{eqn:fcn_prox}
% \end{equation}
% The gradient of \eqref{eqn:fcn_prox} is given by
% \begin{equation*}
%   \nabla f_{\mu}(x)[i] = Wu_{\mu}(x)
% \end{equation*}
% where
% \begin{equation*}
%   (u_{\mu}(x))[i] = 
% \begin{cases}
%     \mu^{-1}(W^Tx)[i], & |W^T x[i]| < \mu,\\
%     \text{sgn}(W^T x[i]), & \text{else}.
%   \end{cases}
% \end{equation*}
% The optimization \eqref{op:bp_tv} may now be solved using Algorithm~\ref{ag:NESTA}.

@@

%   \Initialize{$x \in Q_p, \mu=0.9||W^T x_0||_{\infty}$, $L_{\mu} = \frac{1}{\mu}||W||^2_2$}
%     % $x^{k+1} = x^{k}$\\
%   \While{$\mu < \mu_{\textrm{final}}$ } {
%     $k=0$\\
%       \While{$\Delta f_{\mu} > \delta$} {
%         $\alpha_k = \frac{1}{2}(k+1)$ \\
%         $\tau_k = \frac{2}{k+3}$ \\
%         $\displaystyle y_k = \text{arg } \min_{x\in Q_p}
%           \frac{L_{\mu}}{2}||x - x_k||_2^2 + \langle \nabla f(x_k),~ x\rangle$\\
%         $\displaystyle {z_k = \text{arg } \min_{x\in Q_p} \frac{L_{\mu}}{2}||x-x_0||_2^2 +
%         \langle \sum_{i=0}^k \alpha_i \nabla f(x_i), ~x \rangle}$\\
%         $x_{k+1} = \tau_kz_k + (1-\tau_k)y_k$\\
%         $k \leftarrow k + 1$\\
%       }
%       $x_0 = x_k$\\
%       $\mu = \gamma \mu$\\
%       $L_{\mu} = (1/\mu)||W||^2_2$
%     }
%   \Output{Reconstruction $x_k$}
%   \caption{NESTA algorithm}
%   \label{ag:NESTA}
% \end{algorithm}
%   \todo{seems like a waste of space to reprint the algo, since its really no different than the reference}

% In the termination condition for the inner loop, $\Delta f_{\mu} = |f_{\mu}(x_k) - \bar{f}_{\mu}(x_k)|/\bar{f}_{\mu}(x_k)$ where
% \begin{equation}
% \bar{f}_{\mu}(x_k) = \frac{1}{\min\{10, k\}} \sum_{\ell=1}^{\min\{10,k\}} f_{\mu}(x_{k-\ell})
% \end{equation}
% When $RR^T=I$, the sub-problems in lines 6 and 7 have a closed-form solution. Details can be found in~\cite{becker_nesta_2011}.

% % The sub-problems in lines 6 and 7 have the form
% % \begin{equation}
% %   p = \textrm{arg } \min_{x\in Q_p} \frac{L}{2}||x - r|| + \langle g,~x - r\rangle
% % \end{equation}
% % for some vectors $g$ and $r$ and, as shown in~\cite{becker_nesta_2011}, have a closed-form solution given by
% % \begin{align}
% %   \lambda & = \textrm{max}(0, \sigma^{-1}\left| b - Rq\right| - L\\
% %   q &= r - L^{-1}g\\
% %   p &= \left(I - \frac{\lambda}{\lambda + L} R^TR\right)\left( \frac{\lambda}{L}R^Tb + q\right).
% % \end{align}

@@

Implementing the $\mu$-path scheme involves operating the AFM in several
distinct stages. In the $XY$-direction, the system must transition from tracking
a step command (in the transition to a new measurement location) to tracking a
scan pattern. For our simple integral controller, this only involves changing
the reference signals $X_{ref}$ and $Y_{ref}$.
In the $Z$-direction, the system
must transition between tip descent, surface scanning, tip retraction, and
maintaining a $Z$-axis position.

@@

\begin{table}
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $X$,$Y$ & $Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}
Two cycles of this sequential process are illustrated by the time series in
Fig.~\ref{fig:mupathsignaltrajectory}. The following subsections describe in
more detail several features of the CS scanning process.

\begin{figure}[t!]
  \centering
  \includesvg[width=1\columnwidth]{figures/cs_cycle.svg}
  \caption{Several CS cycles. Each state is indicated by color.}
  \label{fig:mupathsignaltrajectory}
\end{figure}

\subsection{Cantilever does not need to fully disengage.}
\todo{Not clear to me yet that friction is useful if I scan parallel to cantilever.
  Also not clear to me that the normal force and shear force are orthogonal concepts.
  Maybe the macroscopic analogy breaks down, but thinking of a knife pushed into a surface and then dragged...}
In our initial implementation \cite{braker_hardware_2018}, we insisted that the during tip-retraction
(state-5), the cantilever tip should fully break contact with the surface. Here,
we impose no such requirement. Rather, we change the $Z$-setpoint during
tip-retraction to only pull away far enough that the deflection signal stays low
(i.e., below the scanning setpoint), even while we run across the surface.
We choose \rzup so that the interaction between the tip and sample becomes purely adhesive.
In general, stable imaging in is not possible at \rzup and complete dis-engagement occasionally occurs.

The change in setpoints is achieved via the multiplexer in Fig.~\ref{fig:afm_bd_dinv}
which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint
$r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme.

% Ultimately the concern is damage to the sample and tip,
% which is caused by the force imparted by the cantilever probe
% \cite{clayton_review_2009}. The deflection signal, though un-calibrated (meaning
% it does not directly lead to a value in $kN$), is proportional to that force,
% because it is proportional to how much the cantilever is bent. I.e., a large
% (towards $+\infty$) deflection implies the spring force of the cantilever is
% pushing hard into the sample, thereby inducing damage. In a typical imaging
% scenario, we choose a deflection setpoint, (say $0.15$~[v], as in
% Fig.~\ref{fig:z_drift}), and have in some way decided that this is sufficient,
% all else being equal, to not damage things too much.

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1\textwidth]{figures/AFM_loop_xyz.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup is active
    for decision index $i\in\{0,1,5\}$, which correspond to states such that
    $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
\subsection{The pre-scan}
% \begin{figure*}
%   \begin{subfigure}{.33\textwidth}
%     \includesvg[width=1\textwidth]{figures/zbounce_uz_example.svg}
%     \caption{ }
%     % \label{fig: }
%   \end{subfigure}
%   \begin{subfigure}{.33\textwidth}
%     \includesvg[width=1\textwidth]{figures/prescan_uz_example.svg}
%     \caption{ }
%     \label{fig:uz_prescan}
%   \end{subfigure}
%   \begin{subfigure}{.33\textwidth}
%     \includesvg[width=1\textwidth]{figures/prescan_zbounce_PSD.svg}
%     \caption{ }
%     % \label{fig: }
%   \end{subfigure}
%   \caption{(a) A ``$Z$-bounce'' experiment where the $XY$-stage is motionless
%     and the cantilever is repeatedly engaged and disengaged. (b) One cycle of a
%     tightly packed set of CS-cycles where the movement between each measurement
%     location is 0. (c) PSDs from the portions of the trajectories which are
%     colored blue and turquoise, averaged over many such cycles, demonstrating
%     that the scan process itself reduces vibration in the $Z$-axis.}
%   \label{fig:prescan_difference}
% \end{figure*}
After the tip-descent, we begin scanning in a ``pre-scan'' phase (state (3)).
During this phase,

>>>>

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. In this
  paper, we describe a hardware implementation and demonstration of
  one such approach based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  % three approaches: (1) inpainting, an interpolation technique that
  % diffuses information from sampled locations into unsampled
  % regions, (2)
  % basis-pursuit, a compressive sensing-based algorithm that seeks to
  % minimize a
  % measure of sparsity in the underlying image, and (3)
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation (BPVV), that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. \todo{Re-write abstract, especially
    conclusions} These experiments demonstrate that at slow scan
  rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction)
  and that at high rates, the undersampling scheme produces higher
  quality images than raster scanning.
\end{abstract}

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}

@@

% % ------------------ TKTKTKT ------------------
% \todo{Strike this?} First, we demonstrate performance using both
% randomly-placed $\mu-$paths and designed $\mu-$paths that are
% optimized to minimize the expected reconstruction error given some
% initial estimate of the frequency structure of the sample. These
% schemes are described in Sec. \ref{sec:samplingScheme}.
% % --------------------------
We also introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec.
\ref{sec:reconstructionMethods}. Section \ref{sec:control} describes
improvements to the control algorithms compared to our preliminary
implementation in ~\cite{braker_hardware_2018}. We then demonstrate
the entire scheme from sampling through reconstruction using
experiments on a commercial AFM (Agilent 5500). The experimental setup
and implementation details are described in Secs.
\ref{sec:experimentalSetup} and \ref{sec:implementation}\footnote{The
  software portion of the implementation and image reconstruction
  described in this paper can be found at
  \url{https://github.com/rabraker/AFM_CS}}. Experimental results are
provided in Sec. \ref{sec:results}. Finally, we provide a few
concluding remarks in Sec. \ref{sec:conclusions}.

@@

Although BPDN is effective in the general setting, using it for
reconstruction from horizontal $\mu$-path samples often yields
artifacts in the vertical direction (that is, the direction orthogonal
to the $\mu$-path scans) leading to strong discontinuities in the
image. These artifacts grow more prominent as the length of the
$\mu$-paths is increased \cite{maxwell2014compressed}, because these
longer paths lead to increased mutual coherence between the
measurement matrix $R$ and the sparsity basis $U$. However, longer
$\mu$-paths means fewer $\mu$-paths and thus shorter scan times.
\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
                                   & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
                                                          & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
                                         & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
                                                                          & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
                                                                          % ------------------------------
                                                                          % Row
                                                                          % 2
                                                                          % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
                                   & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
                                                          & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
                                         & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
                                                                          & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, 40 pixel long $\mu$-paths with 25\%
    sampling. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_v x=\text{vec}(\nabla_v X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Section 6.1,]{becker_nesta_2011}).
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
    %     (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}.

% For completeness, we give an overview here. To that end, we re-write
% the cost $f(x)$ in \eqref{op:bp_tvW} as
% \begin{equation*}
%   f(x) = \max_{u\in Q_d} \left\langle u, W^Tx \right\rangle,\quad Q_d = \{u:~||u||_{\infty} \leq 1\}.
% \end{equation*}

% One of the core features of the NESTA algorithm is to replace the
% non-smooth functional $f(x)$ with a smoothed approximation
% \begin{equation}
%   f_{\mu} = \max_{u\in Q_d} \left\langle u,
%     W^T x \right\rangle - \frac{\mu}{2}||u||_2^2. \label{eqn:fcn_prox}
% \end{equation}
% The gradient of \eqref{eqn:fcn_prox} is given by
% \begin{equation*}
%   \nabla f_{\mu}(x)[i] = Wu_{\mu}(x)
% \end{equation*}
% where
% \begin{equation*}
%   (u_{\mu}(x))[i] = 
% \begin{cases}
%   \mu^{-1}(W^Tx)[i], & |W^T x[i]| < \mu,\\
%   \text{sgn}(W^T x[i]), & \text{else}.
% \end{cases}
% \end{equation*}
% The optimization \eqref{op:bp_tv} may now be solved using
% Algorithm~\ref{ag:NESTA}.

@@

%   \Initialize{$x \in Q_p, \mu=0.9||W^T x_0||_{\infty}$,
%   $L_{\mu} = \frac{1}{\mu}||W||^2_2$}
%     %   $x^{k+1} = x^{k}$\\
%   \While{$\mu < \mu_{\textrm{final}}$ } {
%   $k=0$\\
%   \While{$\Delta f_{\mu} > \delta$} {
%   $\alpha_k = \frac{1}{2}(k+1)$ \\
%   $\tau_k = \frac{2}{k+3}$ \\
%   $\displaystyle y_k = \text{arg } \min_{x\in Q_p}
%   \frac{L_{\mu}}{2}||x - x_k||_2^2 + \langle \nabla f(x_k),~ x\rangle$\\
%   $\displaystyle {z_k = \text{arg } \min_{x\in Q_p}
%   \frac{L_{\mu}}{2}||x-x_0||_2^2 +
%   \langle \sum_{i=0}^k \alpha_i \nabla f(x_i), ~x \rangle}$\\
%   $x_{k+1} = \tau_kz_k + (1-\tau_k)y_k$\\
%   $k \leftarrow k + 1$\\
% }
%   $x_0 = x_k$\\
%   $\mu = \gamma \mu$\\
%   $L_{\mu} = (1/\mu)||W||^2_2$ } \Output{Reconstruction $x_k$}
%   \caption{NESTA algorithm}
%   \label{ag:NESTA}
% \end{algorithm}
% \todo{seems like a waste of space to reprint the algo, since its
% really no different than the reference}

% In the termination condition for the inner loop,
% $\Delta f_{\mu} = |f_{\mu}(x_k) -
% \bar{f}_{\mu}(x_k)|/\bar{f}_{\mu}(x_k)$ where
% \begin{equation}
% \bar{f}_{\mu}(x_k) = \frac{1}{\min\{10, k\}} \sum_{\ell=1}^{\min\{10,k\}} f_{\mu}(x_{k-\ell})
% \end{equation}
% When $RR^T=I$, the sub-problems in lines 6 and 7 have a closed-form
% solution. Details can be found in~\cite{becker_nesta_2011}.

% % The sub-problems in lines 6 and 7 have the form
% % \begin{equation}
% %   p = \textrm{arg } \min_{x\in Q_p} \frac{L}{2}||x - r|| + \langle
% %   g,~x - r\rangle
% % \end{equation}
% % for some vectors $g$ and $r$ and, as shown
% % in~\cite{becker_nesta_2011}, have a closed-form solution given by
% % \begin{align}
% %   \lambda & = \textrm{max}(0, \sigma^{-1}\left| b -
% %     Rq\right| - L\\
% %   q &= r - L^{-1}g\\
% %   p &= \left(I - \frac{\lambda}{\lambda + L}
% %     R^TR\right)\left( \frac{\lambda}{L}R^Tb + q\right).
% % \end{align}

@@

Implementing the $\mu$-path scheme involves operating the AFM in
several distinct stages. In the $XY$-direction, the system transitions
from tracking a step command in the transition to a new
measurement location to tracking a short ramp during the actual $\mu$-path scan.
For our simple integral controller, this only involves changing the reference signals
$X_{ref}$ and $Y_{ref}$. In the $Z$-direction, the system must
transition between tip descent, surface scanning, and tip retraction. F

@@

\begin{table}
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $r_X$,$r_Y$ & $r_Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}
Two cycles of this sequential process are illustrated by the time
series in Fig.~\ref{fig:cs_cycle}. The following
subsections describe in more detail several features of the CS
scanning process.

\subsection{Cantilever does not need to fully disengage.}
\todo{Not clear to me yet that friction is useful if I scan parallel
  to cantilever. Also not clear to me that the normal force and shear
  force are orthogonal concepts. Maybe the macroscopic analogy breaks
  down, but thinking of a knife pushed into a surface and then
  dragged...}
In our initial implementation \cite{braker_hardware_2018}, we insisted that the during
tip-retraction (state-5), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging in is not possible at \rzup and
complete dis-engagement occasionally occurs. This can be seen in bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

Only moving the tip into the adhesive region during the $XY$ move saves considerable time
compared to completely detaching the tip between each $\mu$-path. By eliminating the normal force into the sample, we are able to take dozens of images with a single probe without a noticeable degradation in image quality. However, completely characterizing the damage done to the tip and sample with this scheme compared complete detachment (which can also damage the probe \cite{meyer_atomic_1992}), remains an outstanding goal.

\begin{figure}[t!]
  \centering \includesvg[width=1\columnwidth]{figures/cs_cycle.svg}
  \caption{Several CS cycles. Each state is indicated by color.}
  \label{fig:cs_cycle}
\end{figure}

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1\textwidth]{figures/AFM_loop_xyz.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup
    is active for decision index $i\in\{0,1,5\}$, which correspond to
    states such that $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
\subsection{The pre-scan}
After the tip-descent, we begin scanning in a ``pre-scan'' phase
(state (3)), which is the turquoise portion of the trajectories in Fig.~\ref{fig:cs_cycle}.
There are two reasons for this. First, we have observed that there is less noise in the deflection signal while the $XY$-stage is in motion compared to sitting still. More importantly, it lets us bring the $X$-axis to its scanning velocity at the same time the transients from the $Z$-axis die out. Without this, we must extend the $\mu$-path by a distance equal to the steady-state error following a ramp. For an arbitrary scan speed, this means the $\mu$-path scan would need an extra 
\begin{equation}
  N = \left\lceil\lim_{z\rightarrow 1} \frac{1-H_x}{z-1}\right\rceil
\end{equation}
time steps, where $H_x$ is closed-loop transfer function for the $x$-axis, including $F_X$.
For our system $N=111$. By initiating the scan while the $Z$-axis settles, we eliminate much of
this overhead.

\subsection{Control}\label{sec:control}
% \begin{figure}
%   \centering
%   \includesvg[width=1\columnwidth]{figures/z_control_design.svg}
%   \caption{FRFs of the $Z$-axis. The blue curve is the open loop
%   response from $u_Z$ to deflection. By inverting the bending mode
%   at 215~Hz, we achieve the black curve in closed-loop.}
%   \label{fig:z_control}
% \end{figure}
The control structure for each axis, shown in the block diagram in
Fig.~\ref{fig:afm_bd_dinv}, is similar and essentially consists of a
set of filters in series with a PI controller. The open-loop frequency
response for the $X$ and $Y$ axes is shown as the black curves in Fig.~\ref{fig:mimo_frf_uxuy} and for the $Z$-axis in Fig.~\ref{fig:z_control}. The first resonance of each axis is the bending mode, which
appears at 215~Hz for the $Z$ axis and about 350~Hz for the $X$ and
$Y$ axes.
